---
title: "Editing functions"
author: "Jakub Chromec"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Editing functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document explains how to add, update and delete stuff in Trello. This requires authentication, as explained in the "Authorized access" vignette.

```{r setup}
library(trelloR)
```

Adding resources
----------------

The `post_model()` function allows you to create new resources. At the very least, you only have to specify a type of `model` (eg. "card"). Any other stuff, such as *name*, *decription* or *labels* has to be passed as a named list to the `body` argument.

The following example creates a new card and places it at the bottom of a list (consult the [Trello API reference](https://developers.trello.com/reference) for query parameters):

```r
payload = list(
  name = "My new card!", desc = "An example card.", 
  pos = "bottom",
  idList = list_id # get it by calling get_board_lists()
)
post_model(model = "card", body = payload)
```
When successful, the request returns a named list of values describing the newly created resource, including its ID. This is very useful as you can immediately capture the ID if you need it.

See `?post_model` for additional arguments.

The wrappers for this function include: `?add_board`, `?add_card`, `?add_checklist`, `?add_checkitem`, `?add_comment`, `?add_label`, `?add_list` and `?add_field`.

Updating resources
------------------

The `put_model()` function follows the same logic as the `post_model()`, but can only be called on existing resources.

See `?update_card`, `?update_card_labels`, `?update_card_members`, `?update_checkitem` and `?update_field`.

Deleting resources
------------------

To delete resources, use `delete_model()` and its wrappers: `?delete_card`, `?delete_checklist`, `?delete_checkitem`, `?delete_field` and `?delete_field_option`.

__Caution:__ Deleted resources cannot be brought back!

Built with
----------

```{r}
sessionInfo()
```
