---
title: "Getting data"
author: "Jakub Chromec"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
```

This document explains how to interact with Trello from R. It includes a short intro for people unfamiliar with the Trello API, and a few recipes for how to get data from public boards. If you need access to private boards or need to create/edit boards, check out the other vignettes in this package.

Introduction
------------

Most things in Trello live on a __Board__. A board encapsulates a hierarchy of resources:  __Members, Teams, Lists, Cards, Custom fields__ and __Actions__. Some of the resources may contain additional data (e.g. cards may have labels, due dates etc.) or other nested resources (e.g. boards may have lists, cards, label definitions etc.)

Each resource can have a parent resource (e.g. a board is a parent resource for a card) and child resources (a card can include actions as a child resource). A resource can have more than one parent (a board and a list are both parents to a card).

To access a resource, you need to know its unique ID, or the ID of its parent resource. In some cases (e.g. with boards or cards), you can use the resource URL instead.

Getting data
------------

The following snippet fetches 5 cards from the [trelloR demo board](https://trello.com/b/Pw3EioMM/trellor-r-api-for-trello). This is a public board and so does not require authentication:

```{r setup}
library(trelloR)
```

```{r, results='hide'}
board = "https://trello.com/b/Pw3EioMM/trellor-r-api-for-trello"
param = list(fields = "id,name,idList,labels")
cards = get_board_cards(board, query = param, limit = 5)
```

The result is a data.frame:

``` {r}
cards
```

Nsted resources can be obtained by querying their parent resource. In the example above, we didn't have to query each card individually - instead, we just queried the board resource to get all of its cards.

A general way of retrieving data from a board is to use the function `get_model()` which allows you to specify parameters of the query. However, `trelloR` also includes a number of wrappers for specific resources. For example: 

- use `get_board_cards()` to get cards from a particular board
- use `get_card_members()` to get a list of people assigned to a card
- use `get_board_fields()` to get custom field definitions
- use `get_card_fields()` to get custom field values

For an overview of all available wrappers, call `?get_board`, `?get_card`, `?get_team`, `?get_member`, `?get_list` or `?get_id`.

If there is no wrapper for the query you need to make, you can call `get_model()` directly or use it to make your own. The following example creates a wrapper that fetches all updates to a given card. This is represented by the "updateCard" action type (notice the filter argument):

```{r}
get_card_updates = function(id, ...) {
  get_model(parent = "card", child  = "actions", id = id, filter = "updateCard", ...)
}
```

Such a function can be called the usual way, supplying a card URL or ID:

```{r, results='hide'}
card_id = cards$id[4]
card_updates = get_card_updates(card_id)
```

```{r}
card_updates
```

See `?get_model` and [Trello API reference](https://developers.trello.com/v1.0/reference) to build your own customized wrappers.

Error handling
--------------

If a request fails, the error code is reprinted in the console. Additional server messages are also included to make debugging easier. See the example below with invalid card ID:

```{r, results='hide'}
tryCatch(
  expr  = get_card_actions(id = "ThisIsGoingToFail"),
  error = function(e) message(e$message)
)
```

After it fails, the request is re-send two more times, with random delay before each attempt. If it fails for the 3rd time, the error message is printed on the console and nothing is returned.

Built with
----------

```{r}
sessionInfo()
```
