---
title: "R API for Trello"
author: "Jakub Chromec"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R API for Trello}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
library(httr)
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

This document explains how to use `trelloR` to add, update, remove or retrieve data from Trello via Trello API.

_NOTE. Accessing private boards requires authorization; see **Accessing private boards**._

Getting public data
-------------------

Most things in Trello live on a __Board__. A board encapsulates a hierarchy of "models", i.e. __Members, Teams, Lists, Cards, Labels, Checklists__ and __Actions__. Each model has a parent model (e.g. a board is a parent model for its cards) and child models (a card can include comments as child models).

Models can be accessed recursively, so you can obtain all the actions from all cards by addressing the whole board, rather than going through specific cards one by one.

To access a model, you need to know its unique ID, or the ID of its parent model. In some cases (e.g. with boards or cards), you can use the model URL instead. This is useful because URLs can be found in the address bar of your browser (unlike IDs). You can also obtain model ID by "searching", which you can read more about later.

The following snippet fetches 5 newest cards from the [Trello Development Roadmap](https://trello.com/b/nC8QJJoZ/trello-development-roadmap) (public board, no authentication required):

```{r, results='hide'}
library(trelloR)
roadmap_url = "https://trello.com/b/nC8QJJoZ/trello-development-roadmap"
cards = get_board_cards(roadmap_url, limit = 5)
```

If there are no complaints from the server, a JSON response is received and converted into a `data.frame`. As it often includes many levels of nested JSON structure, the output can be quite rich:

``` {r}
cards
```

###Fetching specific models
`trelloR` includes a number of wrappers for specific models with easy-to-guess names. This eliminates most of the effort and makes the code easier to read. Examples:

- `get_board_cards()` fetches cards from a particular board
- `get_card_members()` fetches a list of people assigned to a card

For an overview of available functions, call `?get_board`, `?get_card`, `?get_team`, `?get_member`, `?get_list` or `?get_id`.

###Custom requests with `get_model()`
All the `get_` functions in this package are wrappers for `get_model()`. You can call it directly if you lack a particular data-fetching option, or you can define a new function by wrapping `get_model()` and providing some default values. 

The following example creates a function that fetches all updates made to a given card (it uses a filter because updates are a type of action, and there's no need to retrieve *all* actions):

```{r}
get_card_updates = function(id, ...) {
    get_model(parent = "card", child  = "actions", id = id, filter = "updateCard", ...)
}
```

Such functions can be called the usual way, supplying a card URL or ID:

```{r, results='hide'}
card_id = cards$id[1]
card_updates = get_card_updates(card_id, limit = 5)
```

```{r}
card_updates
```


For a detailed list of all possible queries, consult the [Trello API reference](https://developers.trello.com/v1.0/reference).

###Additional arguments for `get_model()`

* `token` is necessary for accessing private boards (see __Getting private data__); defaults to `NULL`
* `limit` defaults to 1000 results; set to 0 to obtain *all* results that match the request
* `query` is useful if you need to provide a list of key-value pairs, such as `query = list(key1 = "value1", key2 = "value2")`; these will get appended to the URL request. Setting `filter = "updateCard"` and `limit=10` is equivalent to `query = list(filter = "updateCard", limit = 10)`; if you need more information about query parameters, see [Trello API reference](https://developers.trello.com/v1.0/reference)
* `url` can be used instead of specifying `parent` and `child`

*Deprecated:*

* `paging` is deprecated - use `limit` instead
* `bind.rows` is deprecated

The following fetches 5 newest updates to cards on the public roadmap board.

```{r, results='hide'}
comments = get_model(parent = "board", child = "actions", id = roadmap_url,
                     filter = "commentCard", limit = 5)
```

###Searching for model ID
Sometimes you don't know the ID/URL but you have other information, such as name, description or other text. In such case, you can call `search_model` function which is equivalent to the search field in Trello.

To limit your search to specific models (e.g. only members), call the appropriately named wrapper (searching requires authorization, so you need to supply a token as well - see __Getting private data__):

```{r eval=FALSE}
search_members("Tony Stark", token = my_token)
```

###Error handling
If a request fails because of a client-side or server-side error, the error code is reprinted in the console. Additional server messages are also included to make debugging easier. See the example with invalid card ID below:

```{r, results='hide'}
tryCatch(
    expr  = get_card_actions(id = "I_have_a_bad_feeling_about_this"),
    error = function(e) message(e$message)
)
```

After a server-side error, the request will be re-send two more times, with random delay before each attempt. If it fails for the 3rd time, the error message is printed in the console and nothing is returned.

Accessing private boards
------------------------

Access to private boards requires authorization. This is done by creating a secure token to communicate with the Trello API. 

To create a token, **login** to Trello and visit the [Developer Start Page](https://developers.trello.com/get-started/start-building#connect). Get your developer credentials there, i.e. your "key" and "secret". Then, call the `get_token()` function to create a token for your project. This will also trigger first-time authorization in the browser (you only have to do it once for a given project):

```r
my_token = get_token(key = your_key, secret = your_secret)
```

You will also be offered an option to store the authentication data in your working directory, in a hidden `'.httr-oauth'` file. Make sure you keep it in a **safe, non-shared** location.

Other arguments to `get_token()` include:

* `appname` is used by R to store and manage your authenticated app; you will see your app in the list of connected apps on the settings page in Trello
* `scope` defaults to `"read"`, but you can add `"write"` and `"account"` too
* `expiration` defaults to `"30 days"`, but you can increase it to `"never"` or decrease to `"1day"` or `"1hour"`

###Using the token to retrieve private data

Having obtained the token, you can call `get_my_boards()` to get an overview of your boards. It returns a `data.frame` with board names and IDs related to the user who authorized the app.

```
my_boards = get_my_boards(my_token)
```

For other functions, just feed the token to the `token` argument:

```
cards = get_board_cards(board_url, token = my_token)
```

Write access
------------

In order to add, update and delete models, you need to create a token with writing permissions, e.g.:

```r
my_token = get_token(key = your_key, secret = your_secret, scope=c("read", "write"))
```

This will allow `trelloR` to issue POST, PUT and DELETE http requests.

###Adding models

The `post_model()` function allows you to create new models. At the very least, you have to specify a `model` (eg. "card") and a `token` with permissions to write. To further specify the model to be added, you may have to use the `path` argument. The actual content of the new model (such as *name*, *decription* or *labels*) has to be passed as a named list to the `body` argument.

The following example creates a new card and places it at the bottom of a list:

```
payload = list(
  name = "My new card!", desc = "An example card.", 
  pos = "bottom",
  idList = list_id # get it by calling get_board_lists()
)
post_model(model = "card", body = payload, token = my_token)
```

Consult the [Trello API reference](https://developers.trello.com/v1.0/reference) for element names. Some of the most frequently used POST requests have been implemented as wrappers for this function. See `?add_board`, `?add_card`, `?add_checklist`, `?add_checkitem`, `?add_comment`, `?add_label` and  `?add_list`.

Additional arguments to `post_model()` include:

* `verbose` which, if `TRUE`, provides verbose output for POST requests
* `reponse` to specify the type of response returned; defaults to `"content"` (a named list containing the content of the newly created model), `"status_code"`, `"headers"` or `"response"` in its entirety
* `on.error` to decide what to do when the requests results in an error: issue a message, a warning or stop
* `...` to pass additional arguments to `httr::POST`

###Updating models

The `put_model()` function follows the same logic as the `post_model()` function, but can only be called on existing models. See `?update_card`, `?update_card_labels`, `?update_card_members` and `?update_checkitem`.

###Deleting models

To delete models, use `delete_model()` and its wrappers: `?delete_card`, `?delete_checklist` and `?delete_checkitem`.

__Caution:__ Deleted models cannot be brought back!

###Disclaimer

`trelloR` is not affiliated, associated, authorized, endorsed by or in any way officially connected to Trello, Inc. (www.trello.com).

###Built with

```{r}
sessionInfo()
```
